Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach

### 2013-09-24 inst/doc/nonparametric-bayes.tex
writing

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\DefineShortVerb[commandchars=\\\{\}]{\|}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

The sudden collapse of fisheries and other ecosystems is an increasingly
widespread phenomenon and a pressing concern for ecological management
and conservation. Ecological dyanaics are frequently complex and
difficult to measure, making uncertainty in our understanding a
prediction a persistent challenge to effective management.

Decision-theoretic approaches provide a framework to determine the best
sequence of actions in face of uncertainty, but only when that
uncertainty can be meaningfully quantified {[}@Fischer2009{]}.

Uncertainty enters at many levels management

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Why fisheries
\end{itemize}

The economic value and ecological concern have made marine fisheries the
crucible for much of the founding work (Gordon 1954; Reed 1979; May et
al. 1979; Ludwig and Walters 1982) in managing ecosystems under
uncertainty. Global trends (Worm et al. 2006) and controversy (Hilborn
2007; Worm et al. 2009) have made understanding these challenges all the
more pressing.

@Fischer @Sethi2005, @Peretti, @Munch,

(Classical division)

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Dynamic uncertainty (stochasticity)
\end{itemize}

From \href{}{Clark (1976)} to \href{}{Reed (1979)}

Other sources of uncertainty \href{}{Roughgarden \& Smith 1996},
\href{}{Sethi \emph{et al.} (2005)}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Parametric uncertainty 
\end{itemize}

As the parameter values for these models must be estimated from limited
data,\\there will always be some uncertainty associated with these
values. This uncertainty further compounds the intrinsic variability
introduced by demographic or environmental noise. The degree of
uncertainty in the parameter values can be inferred from the data and
reflected in the estimates of the transition probabilities.

(Mangel 1985; Schapaugh and Tyre 2013)

\href{http://doi.org/10.1016/0304-38008290023-0}{Ludwig \& Walters 1982}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Structural uncertainty 
\end{itemize}

Estimates of parameter uncertainty are only as good as the parametric
models themselves. Often we do not understand the system dynamics well
enough to know if a model provides a good approximation over the
relevant range of states and timescales (criteria that we loosely refer
to as defining the ``right'' or ``true'' model.) So called structural or
model uncertainty is a more difficult problem than parametric
uncertainty. Typical solutions involve either model choice, model
averaging, or introducing yet greater model complexity of which others
may be special cases (model averaging being one such way to construct
such a model) {[}B. K. Williams (2001); Athanassoglou and Xepapadeas
(2012); @Cressie2009{]}. Even setting aside other computational and
statistical concerns (e.g. {[}@Cressie2009{]}), these approaches do not
address our second concern - representing uncertainty outside the
observed data range.

\href{http://dx.doi.org/10.1890/07-0744.1}{@Cressie2009}.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Limits of state space
\end{itemize}

Often the set of feasible states is larger than the range of the
observed states. Though a system can enter a state outside previously
observed values due to stochastic effects alone (such as a particularly
low or high nutrient year), this concern is most keenly felt in
considering the value of actions that have not yet been applied.

Of all sources of uncertainty, this has potentially recieved the least
attention. It has most often appeared in the concept of active learning
or active adaptive managment, (adaptive probing, ) where it may be
valauble to intentionally force a system far from the observed values
even when the expected value such actions is low, as it provides much
faster learning and consequent reduction of model uncertainty that can
allow greater value to be derived later on. Ludwig and Walters (1982)
show that it may be advantageous to fish an unexploited population very
heavily at first to obtain a better estimate of the recruitment rate.
This intuitive strategy when a population is governed by a Ricker or
Beverton-Holt-like dynamic would clearly be disasterous if instead the
dynamics contained an unforseen tipping point. The best way to learn
where the edge lies may be to walk up to it, but it is also the most
dangerous.

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Complex dynamics
\item
  model choice and model averaging approaches
\end{itemize}

@@

\begin{itemize}
\item
  Concerns of tipping points
\item
  Danger of learning
\item
  MAP
\end{itemize}

Decision making under uncertainty is a ubiquitous challenge of natural
resource management and conservation.
\href{http://doi.org/10.1111/faf.12037}{@Glaser2013}

Here we illustrate how a stochastic dynamic programming (SDP) algorithm
can be driven by the predictions from a Gaussian Process (GP). This
provides two distinct advantages compared with contemporary approaches.
First, using a GP sidesteps the need for an accurate model-based
description of the system dynamics (or a model choice or model-averaging
approach which introduces yet more difficulty).\\Second, unlike
parametric models which can only reflect uncertainty in the
corresponding confidence intervals or posterior distributions of their
parameters, the GP approach provides a state-space dependent
representation of uncertainty. This permits a much greater uncertainty
far from the observed data than near the observed data. These features
allow the GP-SDP approach to find robust management solutions in face of
both limited data and no knolwedge of correct model structure.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Note on ``not magic'': honest uncertainty + SDP
\end{itemize}

The idea that any approach can perform well without either having to
know the model or have particularly good data should immediately draw
suspicion. The reader must bear in mind that the strength of our
approach comes not from black-box predictive power from such limited
information, but rather, by providing a more honest expression of
uncertainty outside the observed data without sacrificing the predictive
capacity near the observed data. By coupling this more accurate
description of what is known and unknown to the decision-making under
uncertainty framework provided by stochastic dynamic programming, we are
able to obtain more robust management policies than with common
parameteric modeling approaches.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Note on comparing models (via value function rather than by ``fit'').
  \emph{Do we need this?}
\end{itemize}

The nature of decision-making problems provides a convenient way to
compare models. Rather than compare models in terms of best fit to data
or fret over the appropriate penalty for model complexity, model
performance is defined in the concrete terms of the decision-maker's
objective function, which we will take as given. (Much argument can be
made over the `correct' objective function, e.g.~how to account for the
social value of fish left in the sea vs.~the commercial value of fish
harvested; see Halpern et al. (2013) for further discussion of this
issue. Alternatively, we can always compare model performance across
multiple potential objective functions.) The decision-maker does not
necessarily need a model that provides the best mechanistic
understanding or the best long-term outcome, but rather the one that
best estimates the probabilities of being in different states as a
result of the possible actions.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Note on terminology: ``Non-parametric'' \emph{Do we need this?}
\end{itemize}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Definition
\item
  Previous application
\item
  Why it is particularly suited to these two problems
\item
  (Why this is a novel application thereof)
\end{itemize}

@@

We infer posterior distributions for the parameters of each model in a
Bayesian context using Gibbs sampling (implemented in R \href{}{R} using
jags, \href{}{jags}). We choose uninformative uniform priors for all
parameters (See Appendix, Figures S1-S3, and Table S1, and the R code
provided). One-step-ahead predictions of these model fits are shown in
Figure 1.

@@

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  All models are ``good fits'' to the originally observed data.
\item
  (Simple model choice immediately leads us astray)
\end{itemize}

@@

@@

Eacho of our models $f(S_t)$ must be estimated from training data, which
we simulate from the Allen model with parameters $r = $
\texttt{r p{[}1{]}}, $K =$ \texttt{r p{[}2{]}}, $C =$
\texttt{r p{[}3{]}}, and $\sigma_g =$ \texttt{r sigma\_g} for $T=$

Error in base::parse(text = code, srcfile = NULL) : 1:16: unexpected ','
1: Tobs$ timesteps,
                  ^
\end{verbatim}

r Xo`.

@@

The decision maker typically seeks to determining the course of actions
(also referred to as the policy) that maximizes the expected value of
some objective function such as net present value derived from the
resource over time.\\Though much can be said on how to choose this value
function appropriately (e.g.~see (Halpern et al.
2013)(http://doi.org/10.1073/pnas.1217689110 ``Achieving the triple
bottom line in the face of inherent trade-offs among social equity,
economic return, and conservation.'')) we will assume this is given.
(Nor is this approach necessarily constrained to maximizing the
expectated value of such a function - the decision-theoretic framework
can be adapted to alternatives such as minimizing the maximum cost or
damage that might be incurred; see Polasky et al. (2011)).

@@

\paragraph{Parametric uncertainty}

\paragraph{Structural Uncertainty}

(B. K. Williams 2001; Athanassoglou and Xepapadeas 2012).

@@

Given parameters for the function $f$ and probability distribution $Z$,
along with a given economic model determining the price/profit
$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
and observed stock $X_t$. This problem can be solved exactly for
discretized values of stock $X$ and policy $h$ using stochastic dynamic
programming (SDP) (Mangel 1985). Problems of this sort underpin much
marine fisheries management today.

A crux of this approach is correctly specifying the functional form of
$f$, along with its parameters. The standard approach uses one of a
handful of common parametric models representing the stock-recruitment
relationship, usually after estimating the model parameters from any
available existing data. Uncertainty in the parameter estimates can be
estimated and integrated over to determine the optimal policy under
under uncertainty (Mangel 1985; Schapaugh and Tyre 2013). Uncertainty in
the model structure itself can only be addressed in this approach by
hypothesizing alternative model structures, and then performing some
model choice or model averaging (B. K. Williams 2001; Athanassoglou and
Xepapadeas 2012).

@@

Concerns over the potential for tipping points in ecological dynamics
(Scheffer et al. 2001) highlight the dangers of uncertainty in
ecological management and pose a substantial challenge to existing
decision-theoretic approaches (Brozović and Schlenker 2011). Because
intervention is often too late after a tipping point has been crossed
(but see Hughes et al. (2013)), management is most often concerned with
avoiding potentially catastrophic tipping points before any data is
available at or following a transition that would more clearly reveal
these regime shift dynamics (e.g. Bestelmeyer et al. 2012).

To illustrate the value of the non-parametric Bayesian approach to
management, we focus on example of a system containing such a tipping
point whose dynamics can still be described by a simple, one-dimensional
parametric model.\\We will focus on a simple parametric model for a
single species (derived from fist principles by Allen et al. 2005) as
our underlying ``reality''.

@@

As a low-dimensional system completely described by three parameters,
this scenario should if anything be favorable to a parametric-based
approach. This model contains an Allee effect, or tipping point, below
which the population is not self-sustaining and shrinks to zero
(Courchamp, Berec, and Gascoigne 2008).

@@

We generate initial observational data under the model described in Eq 1
for $T_{\textrm{obs}}=40$ time steps, under a given arbitrary sequence
of harvest intensities, $h_t$. We consider the case in which most of the
data comes from a limited region of state space (e.g.~near a stable
equilibrium), leaving us without observations of the population dynamics
at very low levels which would be useful in discrimating between
recruitment curves {[}@{]} or demonstrating the existence of a tipping
point (Scheffer et al. 2001).\\Using data simulated from a specified
model rather than empirical data permits the comparison against the true
underlying dynamics, setting a bar for the optimal performance possible.

@@

\paragraph{Terminology: ``Non-parametric''}

The term non-parametric to describe modeling approaches such as Gaussian
processes is a common source of confusion, notwithstanding the fact that
the model is still specified by parameters. Some literature has
introduced the term ``semi-parametric'' in tacit acknowledgment of this,
which no doubt only contributes to the confusion. The problem is further
exacerbated by the several meanings assigned to the term in statistics:
(a) that the method does not assume some particular probability
distribution (e.g.~a non-parametric bootstrap) or (b) that the method
does not assume a fixed structure to the model. Our use is that of (b),
in contrast to the classical approaches that do. The difference is
clearest by way of example.

The use of Gaussian process (GP) regression (or ``kriging'' in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems (Kocijan et al. 2005) and
introduced in the ecological modeling and fisheries management by Munch
et al. (2005). An accessible and thorough introduction to the
formulation and use of GPs can be found in Rasmussen and Williams
(2006).

The posterior distribution for the hyper-parameters of the Gaussian
process model are estimated by Metropolis-Hastings algorithm, again with
details and code provided in the Appendix. Rasmussen and Williams (2006)
provides an excellent general introduction to Gaussian Processes and
Munch et al. (2005) first discusses their application in the context of
population dynamics models such as fisheries stock-recruitment
relationships.

@@

\section{References}

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{color}
\usepackage{fancyvrb}
\DefineShortVerb[commandchars=\\\{\}]{\|}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

Decision making under uncertainty is a ubiquitous challenge of natural
resource management and conservation.

The sudden collapse of fisheries and other ecosystems is an increasingly
widespread phenomenon and a pressing concern for ecological management
and conservation. Ecological dynamics are frequently complex and
difficult to measure, making uncertainty in our understanding a
prediction a persistent challenge to effective management.
Decision-theoretic approaches provide a framework to determine the best
sequence of actions in face of uncertainty, but only when that
uncertainty can be meaningfully quantified {[}@Fischer2009{]}.

Uncertainty enters the decision-making process at many levels: intrinsic
stochasticity in biological processes, measurements, and implementation
of policy
\citetext{\citealp[\emph{e.g.}][]{Reed1979}; \citealp{Clark1986}; \citealp[@Roughgarden1997;][]{Sethi2005}},
parameteric uncertainty
\citetext{\citealp[\emph{e.g.}][]{Ludwig1982}; \citealp[@Hilborn1997;
@McAllister1998;][]{Schapaugh2013}}, and model or structural uncertainty
\citep[\emph{e.g.}][]{Williams2001, Cressie2009, Athanassoglou2012}. Of
these, structural uncertainty incorporates the least a priori knowledge
or assumptions and is generally the hardest to quantify. Typical
approaches assume a weak notion of uncertainty where a correct or
reasonable approximation of the dynamics must be identified from among a
handful of alternative models. Here we consider an approach that
addresses uncertainty at each of these levels without assuming the
dynamics follow a particular (i.e.~parametric) structure.

An additional source of uncertainty that has recieved less
attention\footnote{The concept of adaptive probing is one area that has
  explicitly addressed this kind of uncertainty, reaching rather
  opposite conclusions than what we observe here. Adaptive probing
  strategies follow from ``Dual Control'' or ``Active Adaptive
  Management'' approaches (e.g. \citet{Ludwig1982}) that can trade off
  short term utility by choosing actions that can reduce uncertainty.
  Adaptive probing strategies arise when it is valuable to intentionally
  force a system far from the observed values even when the expected
  value such actions is low, as it provides much faster learning and
  consequent reduction of model uncertainty that can allow greater value
  to be derived later on. For instance, \citet{Ludwig1982} show that it
  may be advantageous to fish an unexploited population very heavily at
  first to obtain a better estimate of the recruitment rate. This
  intuitive strategy when a population is governed by a Ricker or
  Beverton-Holt-like dynamic would clearly be disastrous if instead the
  dynamics contained an unforeseen tipping point. The best way to learn
  where the edge lies may be to walk up to it, but it is also the most
  dangerous.} arises when applying a dynamical model outside the range
of data on which it has been estimated. This extrapolation uncertainty
is felt most keenly in decision-theoretic applications, as (a) exploring
the potential action space typically involves considering actions that
may move the system outside the range of observed behavior, and (b)
decision-theoretic alogrithms rely not only on reasonable estimates of
the expected outcomes, but depend on the weights given to all possible
outcomes \citep[\emph{e.g.}][]{Weitzman2013}.

Concerns over the potential for tipping points in ecological dynamics
\citep{Scheffer2001} highlight the dangers of uncertainty in ecological
management and pose a substantial challenge to existing
decision-theoretic approaches \citep{Brozovic2011}. Because intervention
is often too late after a tipping point has been crossed (but see
\citet{Hughes2013}), management is most often concerned with avoiding
potentially catastrophic tipping points before any data is available at
or following a transition that would more clearly reveal these regime
shift dynamics \citep[e.g.][]{Bestelmeyer2012}.

\emph{Map}

Here we illustrate how a stochastic dynamic programming (SDP) algorithm
{[}\citet{Mangel1988}; @Possingham1997, Marescot2013{]} can be driven by
the predictions from a Bayesian non-parametric (BNP) approach
\citep{Munch2005a}.\\This provides two distinct advantages compared with
contemporary approaches. First, using a BNP sidesteps the need for an
accurate model-based description of the system dynamics.\\Second, the
BNP can better reflect uncertainty that arises when extrapolating a
model outside of the data on which it was fit.\\We illustrate that when
the correct model is not known, this latter feature is crucial to
providing a robust decision-theoretic approach in face of substantial
structural uncertainty.

@@

\begin{itemize}
\item
  Complex dynamics \citep{Glaser2013}(http://doi.org/10.1111/faf.12037
  ``Complex dynamics may limit prediction in marine fisheries'')
\item
  model choice and model averaging approaches
\end{itemize}

@@

\begin{itemize}
\item
  Concerns of tipping points
\item
  Danger of learning
\item
  \emph{MAP}
\end{itemize}

This paper represents the first time the SDP decision-making framework
has been used without an a priori model of the underlying dynamics
through the use of the BNP approach.

In contrast to parametric models which can only reflect uncertainty in
parameter estimates, the BNP approach provides a more state-space
dependent representation of uncertainty.

This permits a much greater uncertainty far from the observed data than
near the observed data. These features allow the GP-SDP approach to find
robust management solutions in face of limited data and without
knowledge of the correct model structure.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \emph{Note on ``not magic''}: honest uncertainty + SDP
\end{itemize}

The idea that any approach can perform well without either having to
know the model or have particularly good data should immediately draw
suspicion. The reader must bear in mind that the strength of our
approach comes not from black-box predictive power from such limited
information, but rather, by providing a more honest expression of
uncertainty outside the observed data without sacrificing the predictive
capacity near the observed data. By coupling this more accurate
description of what is known and unknown to the decision-making under
uncertainty framework provided by stochastic dynamic programming, we are
able to obtain more robust management policies than with common
parametric modeling approaches.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Note on: Why fisheries
\end{itemize}

The economic value and ecological concern have made marine fisheries the
crucible for much of the founding work
\citep{Gordon1954, Reed1979, May1979, Ludwig1982} in managing ecosystems
under uncertainty. Global trends \citep{Worm2006} and controversy
\citep{Hilborn2007, Worm2009} have made understanding these challenges
all the more pressing.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \emph{Note on comparing models} (via value function rather than by
  ``fit'').
\end{itemize}

\emph{Do we need this?}

The nature of decision-making problems provides a convenient way to
compare models. Rather than compare models in terms of best fit to data
or fret over the appropriate penalty for model complexity, model
performance is defined in the concrete terms of the decision-maker's
objective function, which we will take as given. (Much argument can be
made over the `correct' objective function, e.g.~how to account for the
social value of fish left in the sea vs.~the commercial value of fish
harvested; see \citet{Halpern2013} for further discussion of this issue.
Alternatively, we can always compare model performance across multiple
potential objective functions.) The decision-maker does not necessarily
need a model that provides the best mechanistic understanding or the
best long-term outcome, but rather the one that best estimates the
probabilities of being in different states as a result of the possible
actions.

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Background on non-parametric modeling.
\end{itemize}

Addressing the difficulty posed by extrapolation without knowing the
true model requires a nonparametric approach to model fitting: one that
does not assume a fixed structure but rather depends on the size of the
data (e.g.~non-parametric regression or a Dirichlet process). This
established terminology is nevertheless unfortunate, as (a) this
approach still involves the estimation of parameters, and (b),
Statisticians use non-parametric to mean both this property (structure
is not fixed by the parameters) and an entirely different (and probably
more familiar) case in which the model does not assume any distribution
(e.g.~non-parametric bootstrap, order statistics). Some literature thus
uses the term semi-parametric, which merely adds ambiguity to the
confusion.

This non-parametric property -- having a structure explicitly dependent
on the data -- is precisely the property that makes this approach
attractive in face of the limited data sampling challenges discussed
above. Having fit a parametric model to some data, the model is
completely described by the values (or posterior distributions) of it's
parameters. The non-parametric model is not captured by its parameter
values or distributions alone. Either the model scales with the
complexity of the data on which it is estimated (e.g.~nonparametric
heirarchical approaches such as the Dirchlet process) or the data points
become themselves part of the model specification, as in the
nonparametric regression used here. we shall see here.

\begin{itemize}
\item
  Definition
\item
  Previous application
\end{itemize}

The use of Gaussian process (GP) regression (or ``kriging'' in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems \citep{Kocijan2005}, and
was first introduced in the context ecological modeling and fisheries
management in \citet{Munch2005}. An accessible and thorough introduction
to the formulation and use of GPs can be found in \citet{Rasmussen2006}.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Why it is particularly suited to these two problems
\item
  (Why this is a novel application thereof)
\end{itemize}

The posterior distribution for the hyper-parameters of the Gaussian
process model are estimated by Metropolis-Hastings algorithm, again with
details and code provided in the Appendix. \citet{Rasmussen2006}
provides an excellent general introduction to Gaussian Processes and
\citet{Munch2005} first discusses their application in the context of
population dynamics models such as fisheries stock-recruitment
relationships.

@@

We infer posterior distributions for the parameters of each model in a
Bayesian context using Gibbs sampling (implemented in R {[}@RTeam{]}
using jags, {[}@R2jags{]}). We choose uninformative uniform priors for
all parameters (See Appendix, Figures S1-S3, and Table S1, and the R
code provided). One-step-ahead predictions of these model fits are shown
in Figure 1.

@@

While it would be straight forward to condition the GP on passing
through the origin (0,0), the estimate shown here is based only on the
observed data. The observed data from which each model is estimated is
also shown. The observations come from only a limited region of state
space corresponding to unharvested or weakly harvested system. No
observations occur at the theoretical optimum harvest rate or near the
tipping point.

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-Figure2.pdf}
\caption{The steady-state optimal policy (infinite boundary) calculated
under each model. Policies are shown in terms of target escapement,
$S_t$, as under models such as this a constant escapement policy is
expected to be optimal \citep{Reed1979}.}
\end{figure}

The resulting optimal management strategy based on each of the inferred
models is shown in Figure 2, against the optimal strategy given the true
underlying dynamics. Policies are shown in terms of target escapement,
$S_t$. Under models such as this a constant escapement policy is
expected to be optimal \citep{Reed1979}, whereby population levels below
a certain size $S$ are unharvested, while above that size the harvest
strategy aims to return the population to $S$, resulting in the
hockey-stick shaped policies shown.

@@

\begin{itemize}
\item
  All models are ``good fits'' to the originally observed data.
\item
  (Simple model choice immediately leads us astray)
\end{itemize}

Though simple mechanistically motivated models offer the greatest
potential to increase our basic understanding of ecological processes
\citep{Cuddington2013, Geritz2012}, such models can be not only
inaccurate but misleading when relied upon in a quantitative decision
making framework.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  We do not know what the correct models are for ecological systems.
\item
  We have limited data from which to estimate the model -- in
  particular, such models may be misleading in predicting the
  probability of outcomes outside the training data.
\end{enumerate}

These aspects are common to many conservation decision making problems,
which thus merit greater use of non-parametric approaches that can best
take advantage of them.

While expressing larger uncertainty outside the observed data, the GP
can also provide a better fit with smaller uncertainty inside the range
of the observed data. This arises from the greater flexibility of the
Gaussian process, which describes a large family of possible curves.
Despite this flexibility, the GP can be described in relatively few
parameters and is thus far less likely to overfit.

@@

This work was partially supported by the Center for Stock Assessment
Research, a partnership between the University of California Santa Cruz
and the Fisheries Ecology Division, Southwest Fisheries Science Center,
Santa Cruz, CA and by NSF~grant EF-0924195 to MM and NSF grant
DBI-1306697 to CB.

@@

Eacho of our models $f(S_t)$ must be estimated from training data, which
we simulate from the Allen model with parameters $r = $
\texttt{r p{[}1{]}}, $K =$ \texttt{r p{[}2{]}}, $C =$
\texttt{r p{[}3{]}}, and $\sigma_g =$ \texttt{r sigma\_g} for $T=$ 40
timesteps, starting at initial condition $X_0 = $ 5.5. The training data
can be seen in Figure 1.

@@

The decision maker typically seeks to determining the course of actions
(also referred to as the policy) that maximizes the expected value of
some objective function such as net present value derived from the
resource over time.\\Though much can be said on how to choose this value
function appropriately (e.g.~see
\citep{Halpern2013}(http://doi.org/10.1073/pnas.1217689110 ``Achieving
the triple bottom line in the face of inherent trade-offs among social
equity, economic return, and conservation.'')) we will assume this is
given. (Nor is this approach necessarily constrained to maximizing the
expectated value of such a function - the decision-theoretic framework
can be adapted to alternatives such as minimizing the maximum cost or
damage that might be incurred; see \citet{Polasky2011}).

@@

\citep{Williams2001, Athanassoglou2012}.

@@

Given parameters for the function $f$ and probability distribution $Z$,
along with a given economic model determining the price/profit
$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
and observed stock $X_t$. This problem can be solved exactly for
discretized values of stock $X$ and policy $h$ using stochastic dynamic
programming (SDP) \citep{Mangel1988}. Problems of this sort underpin
much marine fisheries management today.

A crux of this approach is correctly specifying the functional form of
$f$, along with its parameters. The standard approach uses one of a
handful of common parametric models representing the stock-recruitment
relationship, usually after estimating the model parameters from any
available existing data. Uncertainty in the parameter estimates can be
estimated and integrated over to determine the optimal policy under
under uncertainty \citep{Mangel1988, Schapaugh2013}. Uncertainty in the
model structure itself can only be addressed in this approach by
hypothesizing alternative model structures, and then performing some
model choice or model averaging \citep{Williams2001, Athanassoglou2012}.

@@

To illustrate the value of the non-parametric Bayesian approach to
management, we focus on example of a system containing such a tipping
point whose dynamics can still be described by a simple, one-dimensional
parametric model.\\We will focus on a simple parametric model for a
single species \citep[derived from fist principles by][]{Allen2005a} as
our underlying ``reality''.

@@

As a low-dimensional system completely described by three parameters,
this scenario should if anything be favorable to a parametric-based
approach. This model contains an Allee effect, or tipping point, below
which the population is not self-sustaining and shrinks to zero
\citep{Courchamp2008}.

@@

We generate initial observational data under the model described in Eq 1
for $T_{\textrm{obs}}=40$ time steps, under a given arbitrary sequence
of harvest intensities, $h_t$. We consider the case in which most of the
data comes from a limited region of state space (e.g.~near a stable
equilibrium), leaving us without observations of the population dynamics
at very low levels which would be useful in discrimating between
recruitment curves {[}@{]} or demonstrating the existence of a tipping
point \citep{Scheffer2001}.\\Using data simulated from a specified model
rather than empirical data permits the comparison against the true
underlying dynamics, setting a bar for the optimal performance possible.

@@

@@

\bibliography{nonparametric-bayes}
